HaploQA vs. R/qtl2 Haplotype Reconstructions

### 2022-12-23 haploqa_qtl2_report.Rmd
Updating report

```{r}

library(rstudioapi)
library(qtl2)
library(dplyr)
library(tidyverse)
library(httr)
library(rvest)
library(data.table)
library(qtl2convert)
library(ggrepel)
library(reshape2)
library(stats)
library(ggplot2)
library(stringr)
library(stringi)
library(lsa)
library(readr)

@@

@@

```{r}
gm_res_dir <- file.path(results_dir, 'RDS', 'gm_results.rds')
gm_pct_dir <- file.path(shiny_pct_fp, 'gm_pct_cpmp.csv')
# ### GigaMUGA mutant
gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
url_domain <- 'http://haploqa-dev.jax.org/'
marker_type <- 'GigaMUGA'
html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
summary_df <- summary_df[summary_df$`% No Call` < 10,]

@@

#gigamuga_results <- list()
#gm_pct_res <- list()
skipped_inds <- list()
for (ind in ind_gm) {
  print(ind)
  skip <- FALSE
  tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = T, qtl2_file_gen = T), error = function(e) {skip <<- TRUE})
  if(skip) {
     skipped_inds <- c(skipped_inds, ind)
     next}
  #df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')

  #gm_pct_res[[ind]] <- df
  gigamuga_results[[ind]] <- sample_gm_res
}
gigamuga_results <- readRDS(gm_res_dir)
saveRDS(gigamuga_results, gm_res_dir)
#write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)

for (ind in names(gigamuga_results)) {
  print(ind)
  df <- ind_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
  gm_pct_res[[ind]] <- df
}
```

@@

```{r}

#haplotype_reconstruction_pipeline('BXD', list_pheno, qtl2_file_gen = T, samples_gen = T)

mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
mini_pct_dir <- file.path(shiny_pct_fp, 'mini_pct_comp.csv')

@@

  minimuga_results <- list()
  mini_pct_res <- list()
  for (ind in ind_mini) {
    #ind <- 'JXN'
    print(ind)
    sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = T, qtl2_file_gen = T)
    df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

    minimuga_results[[ind]] <- sample_mini_res
    mini_pct_res[[ind]] <- df
  }

  mini_pct_res <- rbindlist(mini_pct_res)
  saveRDS(minimuga_results, mini_res_dir)
} else {
  minimuga_results <- readRDS(mini_res_dir)
}

@@

```{r}
#source(paste0(root,"/app.R"))
#shinyApp(ui, server)
```

@@

To summarize the above results, for GigaMUGA, we selected 277 samples from DO, 218 samples from CC, 71 samples from F3, 34 samples from F2, and 288 mutant samples. For MiniMUGA, we selected 25 samples. Among which, DO and CC are complex crosses with 8 contributing founders, F3 and F2 are relatively simple two-founder crosses, whereas the number of contributing founders in the mutant crosses samples can vary from 2 to 8. This is to ensure there are sufficient evidences to evaluate the accuracy of haplotype reconstruction results for crosses of all computational sizes and complexity.

@@

```{r}
# filepath to the csvs
# DO
do_pct_df <- fread(file.path(shiny_pct_fp, 'do_truth_comp.csv'))
# CC
cc_pct_df <- fread(file.path(shiny_pct_fp, 'cc_truth_comp.csv'))
# BXD f3
bxd_pct_df <- fread(file.path(shiny_pct_fp, 'bxd_truth_comp.csv'))
# f2
f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
```

@@

ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA haploqa vs qtl2 - percentage of marker discordance') + xlab('model')

@@

```{r}
#gm_pct_temp <- rbindlist(gm_pct_res)
#gm_plot <- gm_pct_temp %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
#plot_df <- melt(gm_plot)
#plot_df$variable <- 'GigaMUGA Mutant'

#ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant - percentage of marker difference between genail and haploqa') + xlab('model')

>>>>

```{r}
library(rstudioapi)
library(qtl2)
library(dplyr)
library(tidyverse)
library(httr)
library(rvest)
library(data.table)
library(qtl2convert)
library(ggrepel)
library(reshape2)
library(stats)
library(ggplot2)
library(stringr)
library(stringi)
library(lsa)
library(readr)

@@

The below two code chunks represent how the pipeline is used for the mutant mice crosses, where the contributing founders vary between crosses. The process involve a certain extent of filtering because some samples within the cross may only have one contributing founder, which does not work for genAIL as the algorithm requires two or more founders to perform haplotype reconstructions, and there is no simple way of identifying how many contributing fouders are present in each individual.

The function below again, saves the RDS files that contain haplotype reconstruction results for all samples in the same RDS directory as above (/results/RDS/), please make sure to remove th existing RDS file if the results need to be regenerated. The haplotype reconstruction results for each individuals can also be found in the directory where the qtl2 input files for read_cross are being stored - which should be the specified 'qtl2_dir' value in the config file.

@@

```{r, eval=F}
gm_res_dir <- file.path(results_dir, 'RDS', 'gm_results.rds')
gm_pct_dir <- file.path(shiny_pct_fp, 'gm_pct_comp.csv')

if (!file.exists(gm_res_dir)) {
  # ### GigaMUGA mutant
  gigamuga_url <- 'http://haploqa-dev.jax.org/tag/Jackson_Lab_Mahaffey_MURGIGV01_20160102_FinalReport.html'
  url_domain <- 'http://haploqa-dev.jax.org/'
  marker_type <- 'GigaMUGA'
  html_table <- read_html(gigamuga_url) %>% html_nodes("a") %>% html_attr("href")
  url_list <- paste0(url_domain, html_table[grepl('/sample', html_table)])
  summary_df <- sample_summary_scrape(read_html(gigamuga_url), url_list, marker_type)
  summary_df$`% No Call` <- as.numeric(gsub("%", "", summary_df$`% No Call`))
  summary_df <- summary_df[summary_df$`% No Call` < 10,]

@@

  gigamuga_results <- list()
  gm_pct_res <- list()
  skipped_inds <- list()
  for (ind in ind_gm) {
    print(ind)
    skip <- FALSE
    tryCatch(sample_gm_res <- sample_haplotype_reconstruction('MURGIGV01', ind, samples_gen = F, qtl2_file_gen = F), error = function(e) {skip <<- TRUE})
    if(skip) {
       skipped_inds <- c(skipped_inds, ind)
       next}
    df <- ind_geno_comp(sample_gm_res, ind, 'MURGIGV01')

    gm_pct_res[[ind]] <- df
    gigamuga_results[[ind]] <- sample_gm_res
  }
  print(paste0('Skipped individuals:', skipped_inds))
  gm_pct_res <- rbindlist(gm_pct_res)
  saveRDS(gigamuga_results, gm_res_dir)
} else {
  gigamuga_results <- readRDS(gm_res_dir)
}

if (!file.exists(gm_pct_dir)) {
  write.csv(gm_pct_res, gm_pct_dir, row.names = F)
} else {
  gm_pct_res <- fread(gm_pct_dir)
}

#write.csv(gm_pct_res, file.path(shiny_pct_fp, 'gm_pct_cpmp.csv'), row.names = F)

for (ind in names(gigamuga_results)[21:100]) {
  print(ind)
  df <- ind_geno_comp(gigamuga_results[[ind]], ind, 'MURGIGV01')
  gm_pct_res[[ind]] <- df
}
```

@@

```{r, eval=F}
mini_res_dir <- file.path(results_dir, 'RDS', 'MiniMUGA_results.rds')
mini_pct_dir <- file.path(shiny_pct_fp, 'mini_pct_comp.csv')

@@

  minimuga_results <- list()
  mini_pct_res <- list()
  skipped_inds <- list()
  for (ind in ind_mini) {
    print(ind)
    skip <- FALSE
    tryCatch(sample_mini_res <- sample_haplotype_reconstruction('MiniMUGA', ind, samples_gen = F, qtl2_file_gen = F), error = function(e) {skip <<- TRUE})
    if(skip) {
       skipped_inds <- c(skipped_inds, ind)
       next}
    df <- ind_geno_comp(sample_mini_res, ind, 'MiniMUGA')

    minimuga_results[[ind]] <- sample_mini_res
    mini_pct_res[[ind]] <- df
  }
  print(paste0('Skipped individuals:', skipped_inds))
  mini_pct_res <- rbindlist(mini_pct_res)
  saveRDS(minimuga_results, mini_res_dir)
} else {
  minimuga_results <- readRDS(mini_res_dir)
}

@@

A dashboard was built in R Shiny to visualize the karyotype plot results of the haplotype reconstructions, the script to the Shiny app is 'app.R' in this directory. The below code chunk should be able to launch the Shiny app.

```{r, eval = F}
source(paste0(root,"/app.R"))
shinyApp(ui, server)
```

@@

To summarize the results, for GigaMUGA, we selected 277 samples from DO, 218 samples from CC, 71 samples from F3, 34 samples from F2, and 288 mutant samples. For MiniMUGA, we selected 25 samples. Among which, DO and CC are complex crosses with 8 contributing founders, F3 and F2 are relatively simple two-founder crosses, whereas the number of contributing founders in the mutant crosses samples can vary from 2 to 8. This is to ensure there are sufficient evidences to evaluate the accuracy of haplotype reconstruction results for crosses of all computational sizes and complexity.

@@

```{r}
# filepath to the csvs
# DO
do_pct_df <- fread(file.path(shiny_pct_fp, 'do_truth_comp.csv'))
# CC
cc_pct_df <- fread(file.path(shiny_pct_fp, 'cc_truth_comp.csv'))
# BXD f3
bxd_pct_df <- fread(file.path(shiny_pct_fp, 'bxd_truth_comp.csv'))
# f2
f2_pct_df <- fread(file.path(shiny_pct_fp, 'f2_truth_comp.csv'))
# GigaMUGA Mutant
gm_pct_res <- fread(file.path(shiny_pct_fp, 'gm_pct_comp.csv'))
# MiniMUGA
mini_pct_res <- fread(file.path(shiny_pct_fp, 'mini_pct_comp.csv'))
```

@@

ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('MiniMUGA haploqa vs genail - % of marker discordance') + xlab('model')

@@

```{r}
gm_plot <- gm_pct_res %>% group_by(sample_id) %>% summarise(qtl2_pct_diff = sum(haplo_diplotype != qtl2_calls)/n()) %>% as.data.frame()
plot_df <- melt(gm_plot)
plot_df$variable <- 'GigaMUGA Mutant'

ggplot(plot_df) + geom_boxplot(aes(x = variable, y = value)) + ggtitle('GigaMUGA Mutant haploqa vs genail - % of marker discordance') + xlab('model')
